Flowchart of Current Algorithm

### 2009-01-20 docs/flowchart/flowchart.tex
removed another marginpar

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
(paper III; eq.~[8])
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We note that while temperature is evolved
when solving these equations, we do not keep the final temperature,
nor do we use it to compute the final change in enthalpy.  Therefore,
$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F .}

@@

\item[spherical:]
\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}

@@

@@

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\}
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (\ref{Linear System for Multigrid})
using multigrid V-cycles with Gauss-Seidel relaxation.

>>>>

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
(paper III; eq.~[8])
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We note that while temperature is evolved
when solving these equations, we do not keep the final temperature,
nor do we use it to compute the final change in enthalpy.  Therefore,
$T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.
\MarginPar{Currently we do update the temperature if {\tt use\_tfromp}=F}

@@

@@

\MarginPar{Should we time-center $p_0$ now that we have split out {\tt advect\_base\_pres}?}

@@

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\}
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip, \label{Linear System for Multigrid}
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (\ref{Linear System for Multigrid})
using multigrid V-cycles with Gauss-Seidel relaxation.
